Concurrent Programming in Harmony

### 2022-01-28 book/paper.tex
describing "save" expression in book

This causes the context of the thread to be saved in <{s}> and
the thread to be no longer running.
Another thread can revive the thread using the <{go}>
\index{go statement}%
statement:

\begin{code}
<{go c r}>
\end{code}

Here <{c}> is a context and <{r}> is a Harmony value.
It causes a thread with context <{c}> to be added to the state that has
just executed the <{stop}>
\index{stop expression}%
expression.  The <{stop}> expression returns the value <{r}>.

@@

\vspace{1em}
\begin{tabular}{|l|l|}
\hline
Boolean & <{False}>, <{True}> \\
\hline
Integer & <{..., -2, -1, 0, 1, 2, ...}> \\
\hline
String & <{"example"}>, <{.example}> \\
\hline
Program Counter & (method names are program counter constants) \\
\hline
Dictionary & <{{ .account: 12345, .valid: False }}> \\
\hline
Set & <{{}, { 1, 2, 3 }, { False, .id, 3 }}> \\
\hline
Address & <{?lock, ?flags[2], None}> \\
\hline
Context & (generated by <{stop}> expression) \\
\hline
\end{tabular}
\vspace{1em}

@@

A context value captures the state of a thread.
A context is itself composed over various Harmony values.

@@

A thread can call <{get_context()}> to retrieve its current
context.

@@

The statement <{go c e}> starts a thread with context <{c}> that
has executed a <{stop}> expression.  The <{stop}> expression
returns value <{e}>.
The same context can be started multiple times, allowing threads to
\emph{fork}.

>>>>

This causes the context of the thread to be saved in <{!s}> and
the thread to be no longer running.
Another thread can revive the thread using the <{go}>
\index{go statement}%
statement:

\begin{code}
<{go s r}>
\end{code}

Here <{s}> contains the context and <{r}> is a Harmony value.
It causes a thread with the context contained in <{s}> to be added to
the state that has just executed the <{stop s}>
\index{stop expression}%
expression.  The <{stop}> expression returns the value <{r}>.

@@

\vspace{1em}
\begin{tabular}{|l|l|}
\hline
Boolean & <{False}>, <{True}> \\
\hline
Integer & <{..., -2, -1, 0, 1, 2, ...}> \\
\hline
String & <{"example"}>, <{.example}> \\
\hline
Program Counter & (method names are program counter constants) \\
\hline
Dictionary & <{{ .account: 12345, .valid: False }}> \\
\hline
Set & <{{}, { 1, 2, 3 }, { False, .id, 3 }}> \\
\hline
Address & <{?lock, ?flags[2], None}> \\
\hline
Context & (generated by <{stop}> or <{save}> expression) \\
\hline
\end{tabular}
\vspace{1em}

@@

A context value (aka continuation) captures the state of a thread.
A context is itself composed over various Harmony values.

The following operations generate contexts:
\begin{center}
\begin{tabular}{|l|l|}
\hline
<{save e}> & returns a Harmony value (see below) \\
\hline
<{stop p}> & saves context in <{!p}> and stops the thread (see below) \\
\hline
\end{tabular}
\end{center}

The <{save e}> expression, when invoked, returns a tuple <{(e, c)}> where
<{c}> is the context value of the thread right after the <{save}> instruction
finished.
Using <{go c r}> the context can be turned into a new thread that, instead,
returns <{r}> from the <{save e}> expression.  See \autoref{fig:fork}
for an example of how this could be used to \emph{fork} a thread into
two threads.

The <{stop p}> expression stores the context of the thread right after
the expression in <{!p}> (i.e., <{p}> must be an address value) and
terminates the thread.  The thread can later be reinstantiated with
<{go !p r}>, in which case the <{stop}> expression returns <{r}>.
A thread can be for ever suspended using <{stop None}>.

@@

@@

\begin{figure}
\begin{code}
<{:inline:}>
def fork():
    atomically:
        let (r, ctx) = save True:
            result = r
            if r:
                go ctx (False, None)

def main():
    if fork():
        print "parent"
    else:
        print "child"

spawn eternal main()
<{:end:}>
\end{code}
\caption{Using <{save}> and <{go}> to implement \texttt{fork()}}
\label{fig:fork}
\end{figure}

The statement <{go c e}> starts a thread with context <{c}> that
has executed a <{stop}> or <{save}> expression.
The <{stop}> or <{save}> expression returns value <{e}>.
The same context can be started multiple times, allowing threads to
\emph{fork}.  See \autoref{fig:fork} for an example.
